---
title: "Ice Cream Rating Analysis"
author: "Ice Cream Project Team"
date: "3/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 9
```{r include=FALSE}
ratings <- read.csv('data/reviews_updated_v2.csv')
```
We are interested in testing whether the mean rating between ice cream brands is 
equal. To accomplish this, we will use ANOVA in order to test the following 
null hypothesis:

H_0: mu_{bj} == mu_{breyers} == mu_{hd} == mu_{talentyi} \
H_1: mu_{bj} != mu_{breyers} != mu_{hd} != mu_{talentyi}

Average rating and standard deviation for every ice cream brands:
```{r}
with(ratings, tapply(ratings$stars, ratings$brand, mean)) # mean
with(ratings, tapply(ratings$stars, ratings$brand, sd)) # std dev
```

To proceed with the Analysis of variance technique, the following assumptions 
about the underlying distribution of ratings must hold:

1) Independence of observations - this has to be true since the samples 
collected are coming from a vast variety of anonymous users online.
2) Normality - considering the large sample size of observation, normality can 
be assumed through the central limit theorem.
3) Equality of variances - 

```{r}
summary(aov(ratings$stars ~ ratings$brand, data = ratings))
```

The F-statistic is highly significant, hence there is significant evidence NOT 
to reject the null hypothesis that the average rating of ice cream flavor 
is the same amongst all brands.

