---
title: "Ice Cream Analysis"
author: "Anushna Prakash"
date: "3/2/2021"
output: pdf_document
---

```{r, echo = FALSE}
library(Hmisc)
library(dplyr)
setwd("~/UW MS Data Science/BIOST 557 - Applied Stats & Experimental Design/Ice Cream Project")
reviews = read.csv("reviews_with_dummies.csv")
reviews$is_5 = ifelse(reviews$stars == 5, 1, 0)
```

# Introduction
## Summary Stats
Below are the names of the columns in the final processed data set. Most of these are dummy columns developed by the team using Python.

```{r, echo = FALSE}
names(reviews)

# This reduced version uses the binary is_5 for response variable
reviews.reduced = select(reviews, -c(X, ingredients.1, key,
                                     author, date, title,
                                     helpful_yes, helpful_no, text,
                                     taste, ingredients, texture,
                                     likes, name, stars))

# This version uses binary is_5 for response, and removes dummified flavor and brand columns
reviews.binary = select(reviews.reduced, c(1, 2, 70:90, 97))
names(reviews.binary)
```
The stars column contains a rating provided by an online user for a particular ice cream product.

# Tests

## T-Tests

``` {r}
sds = c(sd(reviews$is_5),
        sd(reviews$is_5[reviews$contains_dairy == 0]),
        sd(reviews$is_5[reviews$contains_dairy == 1]))
names(sds) = c("Overall", "Contains Dairy", "Does Not Contain Dairy")
sds
```

Because the standard deviations are different from each other, I will use the Welch T-test to determine if average ratings are different for dairy-containing ice creams from dairy-free ice creams.

$H_0: p_{dairy} = p_{dairy free}$
$\\H_A: p_{dairy} \neq p_{dairy free}$

```{r}
dairy_free = reviews$is_5[reviews$contains_dairy == 0]
dairy = reviews$is_5[reviews$contains_dairy == 1]
prop.test(x = c(sum(dairy_free), sum(dairy)),
          n = c(length(dairy_free), length(dairy)),
          alternative = "two.sided")
```
What about organic vs. non-organic?
```{r}
means = c(mean(reviews$is_5),
        mean(reviews$is_5[reviews$contains_organic == 0]),
        mean(reviews$is_5[reviews$contains_organic == 1]))
names(means) = c("Overall", "Conventional", "Organic")
means

```
$H_0: p_{organic} = p_{conventional}$
$\\H_A: p_{organic} \neq p_{conventional}$


```{r}
non_organic = reviews$is_5[reviews$contains_organic == 0]
organic = reviews$is_5[reviews$contains_organic == 1]
prop.test(x = c(sum(non_organic), sum(organic)),
          n = c(length(non_organic), length(organic)),
          alternative = "two.sided")
```

What about nuts vs. no-nuts?

$H_0: p_{nuts} = p_{nonuts}$
$\\H_A: p_{nuts} \neq p_{nonuts}$


```{r}
nuts = c('almonds', 'peanuts', 'walnuts', 'pecans')
has_nuts = reviews$is_5[(reviews$contains_almonds == 1)
                        | (reviews$contains_walnuts == 1)
                        | (reviews$contains_peanuts == 1)
                        | (reviews$contains_pecans == 1)]
no_nuts = reviews$is_5[(reviews$contains_almonds == 0)
                       & (reviews$contains_walnuts == 0)
                       & (reviews$contains_peanuts == 0)
                       & (reviews$contains_pecans == 0)]

prop.test(x = c(sum(has_nuts), sum(no_nuts)),
          n = c(length(has_nuts), length(no_nuts)),
          alternative = "two.sided")
```
What about chocolate vs. no-chocolate?

$H_0: p_{chocolate} = p_{chocolate}$
$\\H_A: p_{chocolate} \neq p_{chocolate}$

```{r}
chocolate_vector = reviews$is_5[reviews$contains_chocolate == 1]
non_choc_vector = reviews$is_5[reviews$contains_chocolate == 0]
prop.test(x = c(sum(chocolate_vector), sum(non_choc_vector)),
          n = c(length(chocolate_vector), length(non_choc_vector)),
          alternative = "two.sided")
```
What about marshmallow vs. no marshmallow?

$H_0: p_{marshmallow} = p_{no \ marshmallow}$
$\\H_A: p_{marshmallow} \neq p_{np \ marshmallow}$

```{r}
marsh_vector = reviews$is_5[reviews$contains_marshmallow == 1]
non_marsh_vector = reviews$is_5[reviews$contains_marshmallow == 0]
prop.test(x = c(sum(marsh_vector), sum(non_marsh_vector)),
          n = c(length(marsh_vector), length(non_marsh_vector)),
          alternative = "two.sided")
```
What about fruits vs. no fruits?

$H_0: p_{fruits} = p_{no \ fruits}$
$\\H_A: p_{fruits} \neq p_{no \ fruits}$

```{r}
contains_fruits = reviews$is_5[(reviews$contains_banana == 1)
                               | (reviews$contains_mangos == 1)
                               # | (reviews$contains_lemon.juice.concentrate == 1)
                               | (reviews$contains_pineapple == 1)
                               | (reviews$contains_raspberries == 1)
                               | (reviews$contains_strawberries == 1)]
no_fruits = reviews$is_5[(reviews$contains_banana == 0)
                               & (reviews$contains_mangos == 0)
                               # & (reviews$contains_lemon.juice.concentrate == 0)
                               & (reviews$contains_pineapple == 0)
                               & (reviews$contains_raspberries == 0)
                               & (reviews$contains_strawberries == 0)]
prop.test(x = c(sum(contains_fruits), sum(no_fruits)),
          n = c(length(contains_fruits), length(no_fruits)),
          alternative = "two.sided")
```

## Question 9 & 10: ANOVA
We are interested in testing whether the mean rating between ice cream brands is equal. To accomplish this, we will use ANOVA in order to test the following null hypothesis:

$H_0: p_{bj} = p_{breyers} = p_{hd} = p_{talenti}$
$\\H_A: p_{bj} \neq p_{breyers} \neq p_{hd} \neq p_{talenti}$

Average rating and standard deviation for every ice cream brands:
```{r}
brand_vector = tapply(reviews$is_5, reviews$brand, sum) # sum of successes
total_vector = tapply(reviews$is_5, reviews$brand, length) # sum of trials
brand_vector
total_vector
```

To proceed with the Analysis of variance technique, the following assumptions about the underlying distribution of ratings must hold:

1) Independence of observations - this has to be true since the samples collected are coming from a vast variety of anonymous users online.
2) Normality - considering the large sample size of observation, normality can be assumed through the central limit theorem.

```{r}
prop.test(x = brand_vector,
          n = total_vector,
          alternative = "two.sided")
```

$H_0: p_{vanilla} = p_{chocolate} = p_{strawberry}$
$\\H_A: p_{vanilla} \neq p_{chocolate} \neq p_{strawberry}$

```{r}
flavors = c('vanilla', 'chocolate', 'strawberry')
reviews_filtered = reviews %>% filter(Flavor %in% flavors)
flavor_vector = tapply(reviews_filtered$is_5, reviews_filtered$Flavor, sum)
total_vector = tapply(reviews_filtered$is_5, reviews_filtered$Flavor, length)
prop.test(x = flavor_vector,
          n = total_vector,
          alternative = "two.sided")

```


## Linear Regression and Logistic Regression

Linear Regression Assumptions:
1. Indepdence
2. Model is linear
3. Homoskedastic errors
4. No multicollinearity

$$ {star \ rating} = \beta_0 + ...$$

Logistic Regression Assumptions:
1. Independence
2. Response is binary
3. No multicollinearity
4. Linear relationship of betas to log-odds
5. Large sample size

$$ log(\frac{p_{5 \ star}}{1 - p_{5 \ star}}) = \beta_0+\beta_1{bj}+\beta_2{dreyers}+\beta_3{talenti}+\beta_4{vanilla}+\beta_5{chocolate}+...\\ +\beta_{n-}{contains \ chocolate}+...+\beta_{n-1}{contains \ dairy}+\beta_n{organic}$$