---
title: "Ice Cream Analysis"
author: "Anushna Prakash"
date: "3/2/2021"
output: pdf_document
---

```{r, echo = FALSE}
library(Hmisc)
library(dplyr)
setwd("~/UW MS Data Science/BIOST 557 - Applied Stats & Experimental Design/Ice Cream Project")
reviews = read.csv("reviews_with_dummies.csv")

```

# Introduction
## Summary Stats
Below are the names of the columns in the final processed data set. Most of these are dummy columns developed by the team using Python.

```{r, echo = FALSE}
names(reviews)
```
The stars column contains a rating provided by an online user for a particular ice cream product.
```{r, echo = FALSE}
describe(reviews$stars)
hist(reviews$stars, main = "Histogram of Star Ratings", xlab = "Star Rating")
```
# Tests

## T-Tests

``` {r}
sds = c(sd(reviews$stars),
        sd(reviews$stars[reviews$contains_dairy == 0]),
        sd(reviews$stars[reviews$contains_dairy == 1]))
names(sds) = c("Overall", "Contains Dairy", "Does Not Contain Dairy")
sds
```

Because the standard deviations are different from each other, I will use the Welch T-test to determine if average ratings are different for dairy-containing ice creams from dairy-free ice creams.

$H_0: \mu_{dairy} = \mu_{dairy free}$
$\\H_A: \mu_{dairy} \neq \mu_{dairy free}$

```{r}
dairy_free = reviews$stars[reviews$contains_dairy == 0]
dairy = reviews$stars[reviews$contains_dairy == 1]
t.test(x = dairy, 
       y = dairy_free, 
       alternative = "two.sided", 
       mu = 0, 
       var.equal = FALSE,
       conf.level = 0.95)

```
What about organic vs. non-organic?

$H_0: \mu_{organic} = \mu_{conventional}$
$\\H_A: \mu_{organic} \neq \mu_{conventional}$


```{r}
non_organic = reviews$stars[reviews$contains_organic == 0]
organic = reviews$stars[reviews$contains_organic == 1]
t.test(x = organic, 
       y = non_organic, 
       alternative = "two.sided",
       mu = 0,
       var.equal = FALSE,
       conf.level = 0.95)
```

What about nuts vs. no-nuts?

$H_0: \mu_{nuts} = \mu_{nonuts}$
$\\H_A: \mu_{nuts} \neq \mu_{nonuts}$


```{r}
nuts = c('almonds', 'peanuts', 'walnuts', 'pecans')
has_nuts = reviews$stars[(reviews$contains_almonds == 1)
                         | (reviews$contains_walnuts == 1)
                         | (reviews$contains_peanuts == 1)
                         | (reviews$contains_pecans == 1)]
no_nuts = reviews$stars[(reviews$contains_almonds == 0)
                         & (reviews$contains_walnuts == 0)
                         & (reviews$contains_peanuts == 0)
                         & (reviews$contains_pecans == 0)]

t.test(x = has_nuts, 
       y = no_nuts,
       alternative = "two.sided",
       mu = 0,
       var.equal = FALSE,
       conf.level = 0.95)
```

## Question 9 & 10: ANOVA
We are interested in testing whether the mean rating between ice cream brands is equal. To accomplish this, we will use ANOVA in order to test the following null hypothesis:

$H_0: \mu_{bj} = \mu_{breyers} = \mu_{hd} = \mu_{talenti}$
$\\H_A: \mu_{bj} \neq \mu_{breyers} \neq \mu_{hd} \neq \mu_{talenti}$

Average rating and standard deviation for every ice cream brands:
```{r}
with(reviews, tapply(reviews$stars, reviews$brand, mean)) # mean
with(reviews, tapply(reviews$stars, reviews$brand, sd)) # std dev
```

To proceed with the Analysis of variance technique, the following assumptions about the underlying distribution of ratings must hold:

1) Independence of observations - this has to be true since the samples collected are coming from a vast variety of anonymous users online.
2) Normality - considering the large sample size of observation, normality can be assumed through the central limit theorem.
3) Equality of variances - 

```{r}
summary(aov(stars ~ brand, data = reviews))
```

The F-statistic is highly significant, hence there is significant evidence to reject the null hypothesis that the average rating of ice cream flavor is the same between all brands.


$H_0: \mu_{vanilla} = \mu_{chocolate} = \mu_{strawberry}$
$\\H_A: \mu_{vanilla} \neq \mu_{chocolate} \neq \mu_{strawberry}$

```{r}
flavors = c('vanilla', 'chocolate', 'strawberry')
reviews_filtered = reviews %>% filter(Flavor %in% flavors)
with(reviews_filtered, tapply(reviews_filtered$stars, reviews_filtered$Flavor, mean)) # mean
summary(aov(formula = stars ~ Flavor, data = reviews_filtered))

```